---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(lattice)

setwd("D:\\study\\reproducible research")
dat=read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
steps=aggregate(steps ~ date, dat, sum)
hist(steps$steps, 
     main = "Total number of steps taken each day", 
     xlab = "Steps")

smean1=mean(steps$steps)

smedian1=median(steps$steps)
```
The mean and median of the total number of steps taken per day are `r smean1` and `r smedian1` respectively.

## What is the average daily activity pattern?
```{r}
steps1=aggregate(steps ~ interval, dat, mean)
plot(steps1$interval,steps1$steps, type="l", 
     main="Average number of steps per day by interval",
     xlab="Interval", ylab="Number of steps")

smax=steps1$interval[which.max(steps1$steps)]
```

The `r smax` interval contains the maximum number of steps.

## Imputing missing values
```{r}
missing=sum((is.na(dat$steps)))
```

There are `r missing` missing values in the dataset.

```{r}
d1=dat 

for (i in steps1$interval) {
    d1[d1$interval == i & is.na(d1$steps), ]$steps = steps1$steps[steps1$interval == i]
}
head(d1)

steps2=aggregate(steps ~ date, d1, sum)

hist(steps2$steps, 
     main = "Total number of steps taken each day (imputed)", 
     xlab = "Steps")

smean2=mean(d1$steps)

smedian2=median(d1$steps)

smean3=smean1-smean2  # difference in mean before and after imputation 

smedian3=smedian1-smedian2  # difference in median before and after imputation 

sum(is.na(d1))
```

The mean and median of the total number of steps taken per day are `r smean2` and `r smedian2` respectively.
The mean and median changed `r smean3` and `r smedian3` respectively.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
d1$week=weekdays(as.Date(d1$date))
d1$week[d1$week %in% c("Saturday","Sunday")]="weekend"
d1$week[!d1$week=="weekend"]="weekday"

steps3=aggregate(steps ~ interval + week, d1, mean)

xyplot(steps ~ interval | week, data = steps3, type="l",  lwd = 2,
        layout = c(1, 2), 
        main="Average number of steps per day by interval",
        xlab="Interval", ylab="Number of steps")
```



